---
title: "Colours"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Colours}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

```{r colour-tiles, include = FALSE, echo = FALSE}

library(ggplot2)
library(dplyr)
library(tibble)
library(tidyr)

dat <-
  enframe(sgplot::sg_colour_palettes) %>%
  unnest(value) %>%
  mutate(name = sprintf("%-10s", name)) %>%
  group_by(name) %>%
  mutate(col = paste0("col", row_number())) %>%
  ungroup() %>%
  mutate(name = factor(name, levels = unique(.$name))) %>%
  complete(name, col, fill = list(value = NA))

col_tiles <- function(colour_data, scale_ratio = 1) {
  ggplot(colour_data, aes(x = .data$col, y = .data$name, fill = .data$value)) +
  geom_tile(color = "white",
            linewidth = 3,
            linetype = 1) +
  scale_fill_identity() +
  scale_y_discrete(limits = rev) +
  theme_void() +
  theme(axis.text.y = element_text(size = 12, hjust = 0)) +
  coord_equal(ratio = scale_ratio)
}

```

The colours in `sgplot` have been selected using the Government Analysis Function [Data visualisation: colours](https://analysisfunction.civilservice.gov.uk/policy-store/data-visualisation-colours-in-charts/) guidance.

## Main Palette

There are three palettes available in the 'main' family. `main` is the default palette for discrete colour scales, however if only two colours are required, `main2` should be used as the dark blue and orange colours have the best colour contrast. Although `main6` is available to use, it should be avoided where possible, as it's generally considered that use of four or more colours in a plot is excessive. Plot redesign should be considered where possible.

```{r main-palettes, echo = FALSE, fig.height = 4}

dat %>%
  filter(grepl("main", name)) %>%
  col_tiles()

```

## Other Palettes

There are two other colour palettes available; `sequential` and `focus`. 

```{r other-palettes, echo = FALSE, fig.height = 3.5}

dat %>%
  filter(!grepl("main", name)) %>%
  col_tiles()

```


Names and hex codes for available colours and palettes can be viewed by running the following code:

```{r view-colours, eval = FALSE}
# View names and hex codes for all colours
sgplot::sg_colour_values

# View names and hex codes for all colour palettes
sgplot::sg_colour_palettes

# View names and hex codes for `main` colour palette
sgplot::sg_colour_palettes[["main"]]
```
